Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-12 vesicle_code/docs/adhesion/adhesion.tex
more minor changes from yny

Finally, to further characterize the rotating dynamics of a doublet in
the fluid trap, we compute and plot the vorticity and total force on a
doublet with identical vesicles with reduced area $\Delta A=0.7$,
Hamaker constant $\mathcal{H} = 0.7$, separation distance $\delta =
0.4$, and extensional rate $\chi = 0.7$ in
Figure~\ref{fig:extensionalVorticity}.  Starting from a vesicle doublet
with an initial inclination angle $\theta=\pi$, the doublet slowly
begins to rotate around $t\sim 250$.  We notice that the net force on
each vesicle is almost anti-parallel to the diverging flow in the
far-field so to balance the hydrodynamic drag force that pulls the
vesicles apart.  Then, as the doublet starts to rotate, each vesicle
experiences a drag force from both the up and down diverging streaming
flows.  For small extensional rate, the rotation will continue until
$\theta = \pi/2$ so the doublet remains perfectly symmetric with respect
to their mid-plane and each vesicle experiences drag forces that cancel
in both directions.  For slightly higher extensional rate, we observe
that the two vesicles slide against each other as the doublet rotates.
The slight sliding motion leads to a transition in the contact region
membrane shape from a bulging profile to an S-shape profile of nearly
uniform height except near the edge of contact region.  This adjustment
greatly reduces the net force on each vesicle, leading to an equilibrium
configuration with an inclination angle $\theta > \pi/2$ as shown in the
left of Figure~\ref{fig:extensionalInclinationAngle}.  Finally, the
vorticity and the flow between the vesicles approaches zero and the
result is a fluid trap containing the doublet.
\begin{figure}[htp]
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity01.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity02.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity03.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity04.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity05.png} \\
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity06.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity07.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity08.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity09.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity10.png}
  \caption{\label{fig:extensionalVorticity} The vorticity and total
  force of two vesicles in a doublet.  The reduced area is $\Delta A =
  0.7$, the separation distance is $\delta = 0.4$, the Hamaker constant
  is $\mathcal{H} = 0.7$, and the extensional rate is $\chi = 0.7$.}
\end{figure}

@@

Flormann {\it et al.} studied the clustering and packing of vesicles in
free suspension~\cite{FlormannAouane2017_SciReports}.  In engineering
applications of vesicle emulsions, smaller vesicles are often enclosed in
a big vesicle, as shown in Figure~\ref{fig:relaxationManyVes}, where
the left panel shows an initial condition of five vesicles suspended
inside of a larger vesicle.  The second panel from the left shows the
membrane configuration and flow streamlines at an intermediate time.
There is no imposed flow, so the dynamics are governed entirely by the
vesicle forces, including adhesion.  In the absence of confinement,
we would expect the vesicles to align themselves as a rouleaux with the
membranes forming a parachute, male-female, or S-shape
configuration~\cite{FlormannAouane2017_SciReports}.  However, because of
the bounding vesicle and the dense packing, the interior vesicles align
themselves differently.  In addition, several parts of the bounding
membrane closely follow the interior membranes in order to minimize the
adhesion energy.   As a result of force balance, the equilibrium shape
of the bounding vesicle is far from a symmetric shape while the interior
smaller vesicles are of similar shape to each other.  We also plot the
tension (third) and adhesion (fourth) of the final equilibrium
configuration.  We remark that the adhesion force is attractive on all
the vesicle membranes while large positive membrane tension is found
only on the bounding vesicle.  In particular, the interior vesicles form
a flat contact region with a large negative tension.  We are now
conducting simulations of such vesicle configuration under a planar
shear flow.

@@

Recently Liu {\em et al.}~\cite{LiuChuNewbyRead2018_bioRxiv} used
immersed boundary simulations to show that, at a separation distance of
tens of nanometers, the thin film between the two membranes facilitate
the coupling between membranes via strong hydrodynamic interactions. In
particular, they demonstrate numerically that the fluctuation in one
membrane is highly correlated to the other membrane without any physical
contact. It is not clear how thermal fluctuations may affect the
hydrodynamics of vesicles under adhesion. For example, does the
fluctuating hydrodynamics in the thin film between two vesicles enhance
adhesion to keep vesicles bound under linear flows?

>>>>

Finally, to further characterize the rotating dynamics of a doublet in
the fluid trap, we compute and plot the vorticity and total force on a
doublet with identical vesicles with reduced area $\Delta A=0.7$,
Hamaker constant $\mathcal{H} = 0.7$, separation distance $\delta =
0.4$, and extensional rate $\chi = 0.7$ in
Figure~\ref{fig:extensionalVorticity}.  Starting from an initial inclination angle $\theta=\pi$, the doublet 
starts to rotate around $t\sim 250$.  We notice that the net force on
each vesicle is almost anti-parallel to the diverging flow in the
far-field so to balance the hydrodynamic drag force that pulls the
vesicles apart.  Then, as the doublet starts to rotate, each vesicle
experiences a drag force from both the up and down diverging streaming
flows.  For small extensional rate, the rotation will continue until
$\theta = \pi/2$ so the doublet remains perfectly symmetric with respect
to their mid-plane and each vesicle experiences drag forces that cancel
in both directions.  For slightly higher extensional rate, we observe
that the two vesicles slide against each other as the doublet rotates.
The slight sliding motion leads to a transition in the contact region
membrane shape from a bulging profile to an S-shape profile of nearly
uniform height except near the edge of contact region.  This adjustment
greatly reduces the net force on each vesicle, leading to an equilibrium
configuration with an inclination angle $\theta > \pi/2$ as shown in the
left of Figure~\ref{fig:extensionalInclinationAngle}.  Finally, the
vorticity and the flow between the vesicles approaches zero and the
result is a fluid trap containing the doublet.
\begin{figure}[htp]
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity01.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity02.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity03.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity04.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity05.png} \\
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity06.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity07.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity08.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity09.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity10.png}
  \caption{\label{fig:extensionalVorticity} The vorticity and total
  force of two vesicles in a doublet.  The reduced area is $\Delta A =
  0.7$, the separation distance is $\delta = 0.4$, the Hamaker constant
  is $\mathcal{H} = 0.7$, and the extensional rate is $\chi = 0.7$.}
\end{figure}

@@

Flormann {\it et al.} studied the clustering and packing of vesicles in
free suspension~\cite{FlormannAouane2017_SciReports}.  In engineering
applications of vesicle emulsions, smaller vesicles are often enclosed in
a big vesicle, as shown in Figure~\ref{fig:relaxationManyVes}, where
the left panel shows an initial condition of five vesicles suspended
inside of a larger vesicle.  The second panel from the left shows the
membrane configuration and flow streamlines at an intermediate time.
There is no imposed flow, so the dynamics is governed entirely by the
vesicle forces, including adhesion.  In the absence of confinement,
we would expect the vesicles to align themselves as a rouleaux with the
membranes forming a parachute, male-female, or S-shape
configuration~\cite{FlormannAouane2017_SciReports}.  However, because of
the bounding vesicle and the dense packing, the interior vesicles align
themselves differently.  
The bounding membrane evolves dynamically together with all the interior 
vesicles to minimize the total membrane energy.
%In addition, several parts of the bounding
%membrane closely follow the interior membranes in order to minimize the
%adhesion energy.   
%

As a result of force balance, the equilibrium shape
of the bounding vesicle is far from a symmetric shape while the interior
smaller vesicles are of similar shape to each other.  
We also plot the
tension (third) and adhesion (fourth) of the final equilibrium
configuration.  We remark that the adhesion force is attractive on all
the vesicle membranes while large positive membrane tension is found
only on the bounding vesicle.  In particular, the interior vesicles form
a flat contact region with a large negative tension.  We are now
conducting simulations of such vesicle configuration under a planar
shear flow.

@@

Finally, it is not clear how thermal fluctuations may affect the
hydrodynamics of vesicles under adhesion. For example, does the
fluctuating hydrodynamics in the thin film between two vesicles enhance
adhesion to keep vesicles bound under linear flows?
Recently Liu {\em et al.}~\cite{LiuChuNewbyRead2018_bioRxiv} used
immersed boundary simulations to show that, at a separation distance of
tens of nanometers, the thin film between the two membranes facilitate
the coupling between membranes via strong hydrodynamic interactions. In
particular, they demonstrate numerically that the fluctuation in one
membrane is highly correlated to the other membrane without any physical
contact. We are actively pursuing this direction with ongoing modeling and simulations
of adhesive membranes with fluctuating hydrodynamics.
